---
title: "GDP Deflator Development 2000-2022"
author: "Daya Ha-Thanh-Van Dau"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The graph is used for the M-SD 1 7102: Development Economics at Hochschule Rhein-Waal. Therefore the project cannot be publicly on GitHub. If you want to re-run the code, please first download the Excel file: **historical_classification_by_income.xlsx** from the [**repository**](https://gitfront.io/r/dayadau/eFPFzX3NCKxD/gdp-defl-develop-income-level/).

It aims to observe the differences in GDP deflation development grouped by income level in 2000. That is, freeze the development of the countries and observe only the GDP deflation development.

More projects at [**GitHub Profile**](https://github.com/dayadau)

Detailed Portfolio on [**Notion Website**](https://www.notion.so/dayadau/Daya-Ha-Thanh-Van-Dau-7f4cdf7c20ad48619025e7ee64d683b6?pvs=4)

# Code with explanation
## Load libraries
```{r}
library(dplyr)
library(readr)
library(WDI)
library(tidyr)
library(ggplot2)
library(readxl)
library(cellranger)
library(simputation)
library(imputeTS)
library(tinytex)
library(pandoc)
```

## Prepare income class 2000 classification
The classification data is taken out from Excel sheet provided by World Bank DataBank. 

```{r}
Inc2000 <- read_excel("historical_classification_by_income.xlsx", "Country Analytical History") 

income2000 <- Inc2000[c(1,11:238),c(1,2,16)]

income2000 <- income2000 %>% 
  rename(incomeold = ...16) %>%
  rename(iso3c = ...1) %>%
  filter(incomeold!="..")

income2000$incomeold <- factor(
  income2000$incomeold, 
  levels = c("L", "LM", "UM", "H"), 
  labels = c("Low income (2000)", 
             "Lower middle income (2000)", 
             "Upper middle income (2000)", 
             "High income (2000)"))
```

## Load data
Access to WDI databank for most updated data.
```{r}
if(!exists("WDI_df")) {
  WDI_df <- WDI(indicator = c("NY.GDP.DEFL.KD.ZG"),
              start = 2000,
              end = 2022,
              extra = TRUE)
}
```

Rename and only select needed attributes.
```{r}
gdpdfl_2000 <- WDI_df %>%
  rename(GDPdefl = NY.GDP.DEFL.KD.ZG) %>% 
  select(country, iso3c, year, GDPdefl, income) %>% 
  subset(income != "Aggregates" & !(iso3c == "COD" & year == "2000" ))

```

## Data Processing
First will observe the missing value status of the dataset. Then decided how to filter out data then use linear regression to impute.

### Filter out NA
```{r}
# income2000 is the data with the old income category
merge_gdpdefl <- merge(gdpdfl_2000, income2000, by="iso3c", all.x=FALSE)

# Check for missing values
statsNA(merge_gdpdefl$GDPdefl)

# Create group of countries only have 0 or 1 observations
many_na_countries_2000 <- merge_gdpdefl %>%
  filter(is.na(GDPdefl)) %>%
  group_by(country) %>%
  summarise(n()) %>%
  filter(`n()` >= 23)

# Filter out them from the main dataset
many_na_countries_list_2000 <- many_na_countries_2000$country

dataGDPdeflator_2000 <- merge_gdpdefl %>%
  mutate(drop = ifelse(country %in% many_na_countries_list_2000, T, F)) %>%
  filter(drop == F) %>%
  select(-drop)

statsNA(dataGDPdeflator_2000$GDPdefl)

```

### Imputation for NA
```{r} 
# Linear Regression approach

simpdataGDPdeflator_2000 <- impute_lm(dataGDPdeflator_2000, GDPdefl ~ year*country) 

## After imputation
ggplot(simpdataGDPdeflator_2000, aes(x = year, y = GDPdefl, color = country)) +
   geom_line(stat = "identity", show.legend = F) 

summary(simpdataGDPdeflator_2000)

```

## Visualisation
After process the dataset, first will transform by building mean then visualisa in the line graph.
```{r pressure, echo=FALSE}
# Building Mean
GDPdeflmean_2000 <- simpdataGDPdeflator_2000 %>%
  group_by(year, incomeold) %>%
  summarise(mean_GDPdefl = mean(GDPdefl, na.rm=T))

#Plot
ggplot () + 
  geom_line(data=GDPdeflmean_2000, aes(x=year, y=mean_GDPdefl, color = incomeold), size= 1.2)+
  labs(title="Inflation, GDP deflator")+
  ylab("GDP deflator (annual %)")+xlim(2000, 2022)+ylim(-5,30)+
  xlab("Year")+
  theme_bw()+
  scale_colour_brewer (breaks=c("Low income (2000)", "Lower middle income (2000)", "Upper middle income (2000)", "High income (2000)"), palette = "Spectral")

```


